---
title: "607Week9"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(jsonlite)
library(knitr)
```

# Assignment Web APIs 
The New York Times web site provides a rich set of APIs, as described here:  http://developer.nytimes.com/docs You'll need to start by signing up for an API key. Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it to an R data frame. 


#Obtaining Data from NY Times
After Obtaining the API key I used the fromJSON function from the JsonLite package.This allowed me to extract the top news articles from the NY Times Website.The request downloaded 28 observation with 23 columns.I only extracted 4 columns from the universe that I later renamed.Finally, I use the Kable function to format the data frame into a well formatted table.

```{r }
url = "https://api.nytimes.com/svc/topstories/v2/home.json?sectionresponse_format"
api ="&api-key=31b11f11ba3f431d8f769aed3720d527"

getdata = fromJSON(paste0(url,api))
NYTimesData = data.frame(getdata)
FilterData = NYTimesData[, c("results.title", "results.abstract","results.byline","results.url" )]
colnames( FilterData) = c("Title","Snipet","Author","Link")   
kable(FilterData)
```

