---
title: "607 Week5"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("bitops")
library("RCurl")
library("tidyr")
library("dplyr")
library("ggplot2")
library("cowplot")

```

## Instructions
Create a .CSV file (or optionally, a MySQL database!) that includes all of the information above.  You are encouraged to use a “wide” structure similar to how the information appears above, so that you can practice tidying and transformations as described below. 

```{r cars}
url = "https://raw.githubusercontent.com/chrisestevez/DataAnalytics/master/607Week5/arrivalDelays.csv"
Rdata = getURL(url)
MyData = read.csv(text = Rdata,header = TRUE,stringsAsFactors = F,sep=",")

colnames(MyData)[which(names(MyData) == "X")] <- "Airline"
colnames(MyData)[which(names(MyData) == "X.1")] <- "Status"

MyData =MyData %>% drop_na()
MyData[2,1]= "ALASKA"
MyData[4,1]="AM WEST"

MyData
```

## Data transformation
```{r}
FinalData = MyData %>% gather("Airport","Values",3:7) %>%
 spread("Status","Values")
FinalData
```
## Data Summary
In order to effectively compare Alaska and AM West carriers, I analyzed their performance based on percentages in various categories.
```{r}
FinalData= FinalData %>%
 group_by(Airline) %>% 
 summarise(ontimeTotal =sum(`on time`),delayedTotal =sum(delayed), TotalFlights =sum(`on time`+delayed),OnTimep = round( sum(`on time`) / sum(delayed +`on time`), digits=2),Delayp = round(sum(delayed) / sum(delayed +`on time`),digits=2) )
  

FinalData
```
The summary indicates that AM West is is above Alaska airlines in on time flights by 2 percentage points.
```{r,echo=FALSE}

onTime = ggplot(FinalData, aes(Airline, OnTimep,fill = Airline) ) +
  geom_bar(stat="identity", width = 0.5)+
    geom_text(aes(label =   OnTimep), position = position_dodge(width=0.75), vjust = -0.25)+
  ggtitle("Total Ontime %")+
 theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
  

delayChart = ggplot(FinalData, aes(Airline, Delayp,fill = Airline) ) +
  geom_bar(stat="identity", width = 0.5)+
    geom_text(aes(label =   Delayp), position = position_dodge(width=0.75), vjust = -0.25)+
  ggtitle("Total Delay %")+
   theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

plot_grid(onTime, delayChart, align='h')
```

```{r}

airportData = MyData %>% 
  gather("Airport","Values",3:7) %>%
 spread("Status","Values") %>% 
group_by(Airline,Airport) %>% 
 summarise(OnTimep = round( sum(`on time`) / sum(delayed +`on time`), digits=2),Delayp = round(sum(delayed) / sum(delayed +`on time`),digits=2) )
airportData
```
In the chart below we can effectively see that Alaska airline has a better on time performance than AM West across all airports. Therefore, contradicting the previous summary.
```{r,echo=FALSE}
ggplot(airportData, aes(Airport, OnTimep,fill = Airline) ) +
  geom_bar(stat="identity", width = 0.75, position = 'dodge')+
    geom_text(aes(label =   OnTimep), position = position_dodge(width=0.75), vjust = -0.25)+
  ggtitle("Airport Ontime %")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
```
 The chart illustrates a high volume of delays from AM West in San Francisco and Seattle
  
```{r,echo=FALSE}
ggplot(airportData, aes(Airport, Delayp,fill = Airline) ) +
  geom_bar(stat="identity", width = 0.75, position = 'dodge')+
  geom_text(aes(label =   Delayp), position = position_dodge(width=0.75), vjust = -0.25)+
  ggtitle("Airport Delay %")+
   theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))


```
